---
title: "ent_why-models"
format: html
editor: visual
---

```{r}
library(tidyverse)

```

## Why did the training regimen disrupt our measure of routine?

Here, we try and work out why variable training disrupted routines on the doors task.

## Possibilities

We have the following hypotheses about what could have happened:

1.  People executed their context relevant responses in a more variable manner. This predicts that lower routine scores should not be correlated with 'other context' or 'neither context' responses - i.e. these responses should stay low and uniformly distributed.
2.  People in the variable group started anticipating switches to a greater extent, and started testing 'other context' doors the longer they had been in a given context. This predicts that the 'other context' doors are more likely to be selected the longer it has been since the last switch. The key signatures here are that people are not generally confused ('neither context' responses are not included) and the rate of other context responses resets after each switch.
3.  More switches injected more noise into each task representation, causing a general decrement to the quality of the representation over the course of the training session. The probability of 'other context' and 'neither context' responses increases over the course of the training session, without resets.

The goal is to start with the simplest models possible to encapsulate these three hypotheses.

## Hyp 1: More variable context relevant behaviours

```{r}

x <- 1:320 # number of trials
k <- 50 # error rate multiplier (free parameter)
y <- rep(1/(tail(x,1)-head(x,1)), times=length(x))*k

plot(x, y, type="l", col="coral2", ylim=c(0,1), xlab = "trial", ylab="p(oc,nc)")

```

## Hyp 2: Repeating cycle of context switch anticipation

```{r}

x <- x
# am going to read in an example of context switches just for ease
tmp <- read.csv("/home/kelly/Documents/projects/doors-data/data-wrangled/exp_lt_evt.csv")
tmp %>% filter(sub == 3 & ses == 2)
```
